Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Spiral Turbulence

### 2022-08-15 paper/paper.tex

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda$. 
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. On the left,  $\Reyn_i = 900$ is above the linear stability boundary, while the right shows $\Reyn_i = 700$, below the stability boundary. Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$. For $\Reyn_i = 700$, we also use $\Lambda = 20, 30$ At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. }
    \label{fig:ke_vs_t_rei900}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches is closests approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

This is notable for a few reasons. At $\Reyn_i=900, \Reyn_o=-3398$ the system is not in a hysteretic region of parameter space. Thus, we do not expect this behaviour to be due to GQL picking up other, coexisting solutions. 
It also highlights the delicate balance an intermittent laminar/turbulent solution represents. The turbulence in this solution has a very broad range of scales, as evidenced by figure~\ref{fig:urms_tz_rei700}. Understanding how it exchanges energy with the mean flow is crucial to understanding both the saturation of the underlying linear instability as well as the maintenance of the non-linear state. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/rei900_story.pdf}
    \caption{Kinetic energy (left) and flow morphologies (right) for $\Reyn_i = 900$, $\Lambda = 5$ covering one full oscillation cycle. The panels on the right show three times corresponding to the vertical lines in the kinetic energy plot on the left. At the kinetic energy minimum (point 1), the flow shows a IPS-like morphology, while an INT burst occurs at the knee (point 2), and a more space-filling turbulent state occurs at maximum energy (point 3).}
    \label{fig:rei900_lambda5_story}
\end{figure}
Finally, in figure~\ref{fig:rei900_vmean}, we show the azimuthal velocity averaged over $theta$ and $z$, giving the mean deviation from CCF as a function of GQL cutoff $\Lambda$. Note that no conditional averaging \cite{2009PhRvE..80f7301D} is performed here, so this represents an average over both laminar and turbulent regions for DNS. As $\Lambda$ is increased, the profile approaches the DNS case; this is typical of GQL behaviour (figure 7 of \cite{2017JFM...810..412T} shows the same for rotating plane Couette flow). 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_vmean_profile.pdf}
    \caption{Mean deviation from CCF for $\Reyn_i = 900$, $\Reyn_o=-3398$. The heavy line is DNS, the orange line is QL, and all others are GQL with $\Lambda = 1,3,5,10$.}
    \label{fig:rei900_vmean}
\end{figure}

\subsection{Subcritical spiral turbulence}
In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$. Each is initialised with the last timestep of the prior run and run for one viscous time. Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study. HThe right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$. Our very preliminary data for $\Lambda = 30$ show that it recovers from the decay and begins to climb in energy again, similar to the early behaviour from the supercritical runs.\footnote{\textbf{Note to reviewers}: this simulation is still running as of the 31 July due date for the initial manuscript. By the time refereeing begins, we will update this plot and section.}

>>>>

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda$. 
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. $\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary. $\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary. Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$. At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state. }
    \label{fig:ke_vs_t_rei900}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches is closests approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

This is notable for a few reasons. At $\Reyn_i=900, \Reyn_o=-3398$ the system is not in a hysteretic region of parameter space. Thus, we do not expect this behaviour to be due to GQL picking up other, coexisting solutions. 
It also highlights the delicate balance an intermittent laminar/turbulent solution represents. The turbulence in this solution has a very broad range of scales, as evidenced by figure~\ref{fig:urms_tz_rei700}. Understanding how it exchanges energy with the mean flow is crucial to understanding both the saturation of the underlying linear instability as well as the maintenance of the non-linear state. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/rei900_story.pdf}
    \caption{Kinetic energy (left) and flow morphologies (right) for $\Reyn_i = 900$, $\Lambda = 5$ covering one full oscillation cycle. The panels on the right show three times corresponding to the vertical lines in the kinetic energy plot on the left. At the kinetic energy minimum (point 1), the flow shows a IPS-like morphology, while an INT burst occurs at the knee (point 2), and a more space-filling turbulent state occurs at maximum energy (point 3).}
    \label{fig:rei900_lambda5_story}
\end{figure}
Finally, in figure~\ref{fig:rei900_vmean}, we show the azimuthal velocity averaged over $\theta$ and $z$, giving the mean deviation from CCF as a function of GQL cutoff $\Lambda$. Note that no conditional averaging \cite{2009PhRvE..80f7301D} is performed here, so this represents an average over both laminar and turbulent regions for DNS. As $\Lambda$ is increased, the profile approaches the DNS case; this is typical of GQL behaviour (figure 7 of \cite{2017JFM...810..412T} shows the same for rotating plane Couette flow). However, unusually, QL performs slightly better than low order GQL with $\Lambda < 5$. 
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_vmean_profile.pdf}
    \caption{Mean deviation from CCF for $\Reyn_i = 900$, $\Reyn_o=-3398$. The heavy line is DNS, the orange line is QL, and all others are GQL with $\Lambda = 1,3,5,10$.}
    \label{fig:rei900_vmean}
\end{figure}

\subsection{Subcritical spiral turbulence}
In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$. Each is initialised with the last timestep of the prior run and run for one viscous time. Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study. The right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$. At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again, peaking in a state .
